Matching Bullets

### 2015-12-07 imaging-paper.Rnw
cleaning up the code and adding a few comments
Rewrite the explanation of loess groove smoothing

\begin{enumerate}
    \item Fix an $x$ coordinate value (Figure~\ref{fig:loess_step12}).
    \item For each y value, compute a rolling average to smooth out any deviations occurring near the minima. (Figure~\ref{fig:loess_step12}).
    \item Determine the peak of the left groove by finding the first value with a negative discretized derivative. Determine the valley of the left groove by finding the first value with a positive discretized derivative after the value of the peak. (Figure~\ref{fig:loess_step34}).
    \item Determine the peak of the right groove by finding the last value with a positive discretized derivative. Determine the valley of the right groove by finding the last value with a negative discretized derivative before the value of the peak. (Figure~\ref{fig:loess_step34}).
    \item Remove all $y$ values left and right of the identified minima in vertical direction (Figure \ref{fig:loess_step34}).
\end{enumerate}

@@

\begin{figure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step12}Steps 1 and 2 of the Loess fit algorithm. For a fixed crosscut (x held constant at 243.75) height values for part 1 of bullet 1 are plotted across different values of y. The height values are smoothed to remove noise in the deviations.}
<<loess_step12, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw()
p1

subdata <- br111_bullet
subdata$smoothed <- c(rep(NA, 10), rollapply(subdata$value, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
subdata$smoothed2 <- c(rep(NA, 10), rollapply(subdata$smoothed, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
p2 <- qplot(y, smoothed2, data = subdata) + theme_bw() + xlim(c(0,2500))

p2
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step34}Steps 3 - 5 of the Loess fitting algorithm. After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside these boundaries are removed (shown in grey)}
<<loess_step34, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
peak_ind <- head(which(diff(subdata$smoothed2) < 0), n = 1)
    groove_ind <- head(which(diff(tail(subdata$smoothed2, n = -(peak_ind + 10))) > 0), n = 1) + peak_ind + 10

    peak_ind2 <- tail(which(diff(subdata$smoothed2) > 0), n = 1)
    groove_ind2 <- tail(which(diff(head(subdata$smoothed2, n = -(length(subdata$smoothed2) - peak_ind2 + 10))) < 0), n = 1)

    p4 <- qplot(subdata$y[subdata$y < subdata$y[groove_ind]], subdata$smoothed2[subdata$y < subdata$y[groove_ind]], colour=I("grey60"), alpha=I(.25)) +
        ylim(c(min(subdata$value[!is.nan(subdata$value)]) - 25, max(subdata$value[!is.nan(subdata$value)]) + 25)) +
        xlim(c(min(subdata$y) - 25, max(subdata$y) + 25)) +
        geom_point(aes(x, y), data = data.frame(x = subdata$y[subdata$y > subdata$y[groove_ind2]], y = subdata$smoothed2[subdata$y > subdata$y[groove_ind2]]), colour=I("grey60"), alpha=I(.25)) +
        geom_point(inherit.aes = FALSE, aes(x, y), data = data.frame(x = subdata$y[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]], y = subdata$smoothed2[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]])) +
        theme_bw() +
        geom_vline(xintercept = subdata$y[peak_ind], colour = "red") +
        geom_vline(xintercept = subdata$y[groove_ind], colour = "blue") +
        geom_vline(xintercept = subdata$y[peak_ind2], colour = "red") +
        geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue")

@@

Things to be dicussed in the future
\begin{enumerate}
\item 
\end{enumerate}
\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{references}

>>>>

\begin{enumerate}
    \item Fix an $x$ coordinate value (Figure~\ref{fig:loess_step1}).
    \item For each $y$ value, compute a rolling average with a particular \textbf{smoothing factor} to smooth out any deviations occurring near the minima. (Figure~\ref{fig:loess_step2}).
    \item For each smoothed $y$ value, compute another rolling average using the same smoothing factor above. (Figure~\ref{fig:loess_step3}).
    \item Determine the peak of the left groove by finding the first doubled-smoothed value $y_i$ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$. Call this peak $p_{left}$. (Figure~\ref{fig:loess_step47}).
    \item Determine the valley of the left groove by finding the first doubled-smoothed $y_j > p_{left}$ such that $y_j < y_{j - 1}$ and $y_j < y_{j + 1}$. Call this valley $v_{left}$.
    \item Reverse the order of the $y$ values and repeat the previous two steps to find the peak and valley of the right groove, $(p_{right}, v_{right})$.
    \item Remove all $y_i < v_{left}$ and $y_i > v_{right}$. (Figure \ref{fig:loess_step47}).
\end{enumerate}

@@

\begin{figure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm. For a fixed crosscut (x held constant at 243.75) height values for part 1 of bullet 1 are plotted across different values of y.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw()
p1
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step2}Step 2 of the Loess fit algorithm. The height values are smoothed with a particular smoothing factor to remove noise in the deviations.}
<<loess_step2, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
subdata <- br111_bullet
smoothfactor <- 35

value_filled <- na.fill(subdata$value, "extend")
smoothed <- rollapply(value_filled, smoothfactor, function(x) mean(x))
smoothed_truefalse <- rollapply(smoothed, smoothfactor, function(x) mean(x))

p2 <- qplot(1:length(smoothed), smoothed) + theme_bw()
p2
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step3}Step 3 of the Loess fit algorithm. The smoothed height values are smoothed one more time with the same smoothing factor.}
<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
p3 <- qplot(1:length(smoothed_truefalse), smoothed_truefalse) + theme_bw()
p3
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step47}Steps 4 - 7 of the Loess fitting algorithm. After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys). Values outside these boundaries are removed (shown in grey)}
<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)

    peak_ind_smoothed <- head(which(rollapply(smoothed_truefalse, 3, function(x) which.max(x) == 2)), n = 1)
    peak_ind <- peak_ind_smoothed + floor(lengthdiff / 2)
    groove_ind <- head(which(rollapply(tail(smoothed_truefalse, n = -peak_ind_smoothed), 3, function(x) which.min(x) == 2)), n = 1) + peak_ind

    peak_ind2_smoothed_temp <- head(which(rollapply(rev(smoothed_truefalse), 3, function(x) which.max(x) == 2)), n = 1)
    peak_ind2_temp <- peak_ind2_smoothed_temp + floor(lengthdiff / 2)
    groove_ind2_temp <- head(which(rollapply(tail(rev(smoothed_truefalse), n = -peak_ind2_smoothed_temp), 3, function(x) which.min(x) == 2)), n = 1) + peak_ind2_temp

    peak_ind2 <- length(subdata$value) - peak_ind2_temp + 1
    groove_ind2 <- length(subdata$value) - groove_ind2_temp + 1

    p4 <- qplot(subdata$y[subdata$y < subdata$y[groove_ind]], subdata$value[subdata$y < subdata$y[groove_ind]], colour=I("grey60"), alpha=I(.25)) +
        ylim(c(min(subdata$value[!is.nan(subdata$value)]) - 25, max(subdata$value[!is.nan(subdata$value)]) + 25)) +
        xlim(c(min(subdata$y) - 25, max(subdata$y) + 25)) +
        geom_point(aes(x, y), data = data.frame(x = subdata$y[subdata$y > subdata$y[groove_ind2]], y = subdata$value[subdata$y > subdata$y[groove_ind2]]), colour=I("grey60"), alpha=I(.25)) +
        geom_point(inherit.aes = FALSE, aes(x, y), data = data.frame(x = subdata$y[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]], y = subdata$value[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]])) +
        theme_bw() +
        geom_vline(xintercept = subdata$y[peak_ind], colour = "red") +
        geom_vline(xintercept = subdata$y[groove_ind], colour = "blue") +
        geom_vline(xintercept = subdata$y[peak_ind2], colour = "red") +
        geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue")

@@

Things to be dicussed in the future
\begin{enumerate}
\item reduce the stretch of crosscuts and match smaller pieces. This allows for both a more realistic scenario, as well as a quantification of the individuality of the segment.
\end{enumerate}
\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{references}
